\bf An HMM-based part-of-speech tagger for Turkish

### 2015-12-08 paper/paper.tex
Comment on some of the code

\documentclass{article}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{hyperref}

@@

\begin{document}
\maketitle

@@

\subsection{Performance}
Parsing the entire corpus and creating a model out of it takes 10.5 seconds on
average. Tagging a sentence (which we're assuming to have 5 words on average)
takes $~0.3$ seconds. The implementation of the \texttt{Data.HMM} datatype uses
\texttt{Data.Array}'s
\footnote{https://hackage.haskell.org/package/array-0.5.1.0/docs/Data-Array.html}
which is why it runs reasonably fast.

>>>>

\documentclass{article}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{mathrsfs}

@@

\begin{document}
%include agda.fmt
\maketitle

@@

\subsection{A Brief Overview on Implementation}
The project consists of two executables: (1) \texttt{create-model} that reads
the data from the corpus and parses it into a form from which we can easily
construct an \texttt{HMM}. Unfortunately, it is highly impractical---if not
impossible---to directly serialize \texttt{HMM} since it has three fields
containing functions
\footnote{http://stackoverflow.com/questions/17785916/can-haskell-functions
  -be-serialized for further information.}. We consider this a fault in the
design of the \texttt{Data.HMM} package and urge further development. Although
this has almost no effect on performance, it obfuscates the code.

Since processing the corpus is a resource-demanding operation,
\texttt{create-model} is intended be run only once after which \texttt{run-tag}
can be used to make use of that data.

\paragraph{\texttt{create-model}}
The executable \texttt{create-model} is implemented in two files:
\texttt{Parse.hs} and \texttt{Main.hs}. \texttt{Parse.hs} handles the parsing of
the corpus.

The function \texttt{freqMap} takes as input a list of orderable elements
(denoted with \texttt{Ord a}) and creates a \texttt{Map} containing, as value,
the frequency of each element.

\begin{lstlisting}[
  language=Haskell,
  mathescape,
  columns=fullflexible,
  basicstyle=\small\ttfamily
]
freqMap :: Ord a $\Rightarrow$ [a] $\rightarrow$ M.Map a Int
freqMap unigrams = populate M.empty unigrams
  where populate :: Ord a M.Map a $Z$ $\rightarrow$ [a] $\rightarrow$ M.Map a Int
        populate m [] = m
        populate m (x:xs) = let freq = (M.findWithDefault 0 x m) :: ℤ
                            in populate (M.insert x (freq + 1) m) xs
\end{lstlisting}

This is the basis of our \emph{supervised} learning.

\paragraph{\texttt{tag}}

The executable \texttt{run-tag} is where the actual tagging happens. The
frequencies we have pre-computed in \texttt{create-model} are passed in to the
probability function which we give below.
\begin{lstlisting}[
  language=Haskell,
  mathescape,
  columns=fullflexible,
  basicstyle=\small\ttfamily
]
probability :: (Ord a, Ord b) $\Rightarrow$ (a, b) $\rightarrow$ Map a Int$\rightarrow$ Map (a, b) ℤ $\rightarrow \mathscr{L}$
probability (x, y) c$\_1$ c$\_2$ = if xCount == 0 || yCount == 0
                           then logFloat $\epsilon$
                           else (logFloat yCount) / (logFloat xCount)
  where yCount = fromIntegral (findWithDefault 0 (x, y) c₂) ∷ Double
        xCount = fromIntegral (findWithDefault 0 x c₁) ∷ Double
\end{lstlisting}
\subsection{Performance}
Parsing the entire corpus and creating a model out of it takes 10.5 seconds on
average. Tagging a sentence (which we're assuming to have 5 words on average)
takes $~0.3$ seconds. The implementation of the \texttt{Data.HMM} datatype uses
\texttt{Data.Array}
\footnote{https://hackage.haskell.org/package/array-0.5.1.0/docs/Data-Array.html}
which is why it runs reasonably fast.
